% Module 02:
  Contour maps and some common quadric surfaces
% Math 251 Calculus 3
% Fall 2013 

```{r echo=FALSE,results="hide",label="options"}
require(knitr, quietly=TRUE)
opts_chunk$set(fig.align="center")
```

```{r echo=FALSE,warning=FALSE,error=FALSE,message=FALSE,results="hide"}
library(mosaic,quietly=TRUE)
trellis.par.set(theme=col.mosaic())
```

Welcome to Module 2. You should have been able to make some progress on 
WeBWorK assignment 1. In this module, we'll discuss the remaining topics 
you'll need to understand to complete it.

### Contour maps

The easiest way to visualize 2-variable functions is with *contour maps*. 

In a contour plot, the points in the plane represent different output values
coming from the various inputs. Here is a contour plot from the [shopping example][shex] in
[Module 01][Module 01]. Shirts are $\$15$ and trousers are $\$20$. The cost of the trip (both ways) is $\$4$.

```{r, linear_contour_plot, fig.align="center"}
linear <- makeFun(4 + 15*x + 20*y ~ x & y)
plotFun(linear(x=x,y=y) ~ x & y, x.lim=c(0,10), y.lim=c(0,10), npts=200, nlevels=10)
```

> Even though the axes are labeled $x$ and $y$, you can infer from the code above the figure which is which. Use the ```xlab=""``` and ```ylab=""``` options to ```plotFun()``` to specify your own labels. 

The points of the plane are grouped into different colors. Each color forms a diagonal line across the plot, called a contour. The colors 
represent the different output values (the labels on the contours) that are possible with our function. 
When two points share a color, they give the same value of the function, and 
when two points have different colors, they give different values of the 
function. 

For each fixed hue, the points that are colored with it form a curve (or 
several curves) in the plane, called a *contour*. Mathematicians also call 
them "level curves". Here is an example of a contour plot of a nonlinear function. What do you notice about this plot that distinguishes it from the first plot?

```{r, quadric_contour_plot}
quad <- makeFun(x^2 - 4*x*y + 2*y^2 ~ x & y)
plotFun(quad(x=x,y=y) ~ x & y, x.lim=c(-6,6), y.lim=c(-3,3), npts=200, nlevels=20)
```

These are easily created in R, as you see above<!-- , and Sage -->. The following R code[^usingR] initializes a random 2-variable function and includes its contour plot into the document, as you see.

```{r, random_contour_plot}
    random <- rfun( ~ x & y, seed = 1066)
    plotFun( random(x=x,y=y) ~ x & y, x.lim=c(-4,4), y.lim=c(-4,4))
```

By default, this produces a full-color contour map. Use the ```plotFun()``` option ```filled=FALSE``` to generate a black and white contour map.

<!-- Sage lacks a facility for machine-generated random functions (as far as I am 
aware), but it can plot a function the user provides quite readily. See the documentation for the ```ContourPlot``` function. 

```{sage}

``` -->

> Like most plotting commands, the R <!-- and Sage  -->commands mentioned here have many options to configure how the plot is generated. You should train yourself in the habit of looking up unfamiliar commands using R Studio's builtin help interface. Try ```?plotFun``` for a start. Notice that the Help pane has Back/Forward buttons and internal history, like a web browser.

**Exercise**. What do you notice about the contours in the contour plot of 
the shopping function?

### Cross-sectional functions

If you read along a horizontal or vertical line in the plane, you can get an 
idea of how changes in either of the inputs affect the output. **MORE IS 
NEEDED**

### Surfaces and graphs

Returning to the abstract geometry of $\mathbf{R}^3$, it is time to consider 
the meaning of equations among the three variables $x$, $y$, and $z$.

It is important to remember that most equations are neither "true" or 
"false" without further context. For example, the equation
$$ a^2 + b^2 = c^2 $$
might be either, for certain $a$, $b$, or $c$, either true or false. Because 
of this, it lacks a truth-value without such context. The equations we graph 
in calculus are similar. It is just this vagueness that allows us to use 
equations to describe curves in the plane.

> An equation involving $x$ and $y$, but no other variables, describes a curve in the plane $\mathbf{R}^2$. A point of the plane is said to lie on this curve when its coordinates, when substituted in the curve's equation, yield truth. In this case, the pair $(x,y)$ is said to *be a solution* of the equation.

Just as above, an equation involving $x$, $y$, and $z$, but no other variables, describes a surface in $\mathbf{R}^3$. A particular point lies on the surface precisely when its coordinates are a solution of the equation.

In [Workshop 00][Workshop 00], you developed formulas giving the distance between any pair of points in $\mathbf{R}^3$ as well as the distance from any point to a coordinate plane or axis. These formulas suffice to construct all spheres, circular cylinders, and circular cones.

> **Exercise**. Remind yourself of formulas for the distance from an arbitrary point $(x,y,z)$ to the coordinate planes and axes. Can you think of a mnemonic trick to remember which is which?

Recall also from [Workshop 00][Workshop 00] that it is often very convenient to deal with *squares* of distances, rather than distances themselves.

The simplest curved surface to describe is a sphere. A sphere is easily described in terms of distance:

> A *sphere* is a closed surface whose interior has a special point, called the *center*, such that all the segments connecting the center with points of the sphere are equal in length. This length is called the *radius* of the sphere.

The definition implies that, for any point $(a,b,c)$ and any radius $r \geq 0$, there is a unique sphere with center $(a,b,c)$ and radius[^degen] $r$. If $(x,y,z)$ is a point on the sphere, the distance from it to the center must be $r$, and this observation yields the classical equation describing the coordinates of points on the sphere.

$$ (x-a)^2 + (y-b)^2 + (z-c)^2 = r^2 $$

As before, working with squares of distances minimizes how much we must put up with unpleasant square roots.

The other surfaces which concern us at present---the cones and the circular cylinders---are also obtained by the equating of various distances. 

> A *circular cylinder* is a surface whose interior has a special line, called the *axis* of the cylinder, such that all points on the surface have the same distance to the axis. To nobody's surprise, this length is called the *radius* of the cylinder.

The best, most special lines in the plane are, of course, the coordinate axes. 

> Use the definition above and the formulas from [Workshop 00][Workshop 00] to write down an equation for the coordinates of points on a cylinder of radius $r$ whose axis is the $z$-axis.

The last class of surfaces we will discuss here is the circular cones[^right].

> A *circular cone with aperture $\pi/2 = \tau/4$* is a surface characterized by a line, called the *axis* of the cone, and a plane perpendicular to the axis. Points on the cone are equidistant from the axis and the plane.

As with cylinders, it is possible to find equations for arbitrary cones: those whose axes and perpendicular planes are completely random. But in general, it is not so useful to do so. It would be smarter to choose coordinates in a clever way, so that it is not even necessary.

> Use the definition above and the formulas from [Workshop 00][Workshop 00] to write down an equation for the coordinates of points on a circular cone.

### Slicing graphs

The idea of "slicing graphs" is one that is probably familiar to you, but with disguised terminology. 

#### The "vertical line test" revisited

"The vertical line test" is a criterion used to decide whether a plane curve is the graph of a function (of the horizontal coordinate). To employ it, you visualize all the vertical lines in the plane. Each of these gives a "slice" of the curve---the intersection of the vertical line with the curve. The vertical line test says that the curve is the graph of a function if, and only if, no slice has more than one point in it.

As soon as you find a slice with two (or more) points in it, then you know the curve is not the graph of any function of the horizontal coordinate. 

Slicing curves in the plane is not so interesting, because most of the time the slices are boring: a handful of points, possibly empty. Think of the vertical lines as the knives doing the slicing. In the cases of interest to us, the knives won't necessarily be lines. Instead, they will be planes.

The knives don't have to be vertical for slicing to make sense. It would be possible to talk about the slice obtained by intersecting the curve with any line, of any slope.

#### Slicing in 3 dimensions

3 dimensions is one up from 2, and *surfaces* are one up from *curves*. So what does it mean to slice a surface? We have to say what the knife is. Usually, it will be some plane. The *slice* we get is the intersection of that plane with the surface---usually, a handful of curves.

> **Exercise**. What are the slices that can result from slicing a sphere?
Make sure you include all the possibilities for what the intersection looks like.

> **Exercise**. What does the slice of a circular cylinder look like if the slicing plane is perpendicular to the cylinder? If it is parallel? If it is neither? What can you say about the shapes that are so obtained?

#### The vertical line test for 2-variable functions

Slicing can be used to determine whether a particular surface is the graph of a function of the two horizontal coordinates. 

> **Exercise**. Describe the knives used to slice for the 3-dimensional vertical line test.

> **Exercise**. Formulate the vertical line test for 2-variable functions.

#### Coordinate slices and level curves

The most important special case of slicing surfaces is when the knife is a plane parallel to one of the coordinate planes. The case of the plane $z = 0$ (that is, the $(x,y)$-plane) merits special attention.

> *Definition*. If $f(x,y)$ is a 2-variable function, the slice of its graph obtained by slicing with the plane $z = c$ is called the level curve of $f(x,y)$ with level $c$. 

<!--document ends; links past here-->

[Module 01]: ../01/Module.html
[shex]: ../01/Module.html#modeling-with-linear-functions
[Workshop 00]: ../../Workshops/00/Workshop.html 

[^degen]: We allow the case $r = 0$, even though the sphere degenerates to a single point. Such degenerate cases are harmless, and often useful.

[^right]: You may be listening, especially if you have taken certain standardized mathematics tests, for the phrases "right circular cylinder" and "right circular cone". The word "perpendicular" in the definition of the circular cone hints where it has gone; in the definition of the cylinder its absence is a different mystery.

[^usingR]: If you are using R Studio installed on your personal computer, you may need to install additional packages, in particular the <pre>mosaic</pre> package. I recommend you use [the College's server](https://rstudio.collegeofidaho.edu/) in preference to a local installation.