\documentclass[12pt]{exam}
\usepackage{cofi}
\usepackage{fourier}
\frenchspacing
\usepackage{parskip}
\firstpageheader{}{}{}
\runningheader{\textbf{Fall 2014}}
 {}
 {\textbf{Math 275}}
 %{\emph{Page \thepage~of \numpages}}
\runningheadrule

\pagestyle{head}
\extrawidth{1in}
\extraheadheight[-0.6in]{-0.1in}
\extrafootheight{-0.5in}
\begin{document}
\noindent
\textbf{{\large Math 275 \hfill Workshop 01}}
% \hfill Name: \underline{\hspace{0.5in}Answers\hspace{2in}}

\noindent
\makebox[\textwidth]{September 10, 2013 \hfill} 

\noindent
\makebox[\textwidth]{Due: Friday, September 12 \hfill Name: \underline{\hspace{3in}}} 

\noindent

\section{Slicing spheres}

This workshop will confirm what is, to varying degrees for various
people, intuitively clear: that all the nonempty slices of a sphere by
planes are circles (we adopt the severely reasonable convention that a
circle of radius 0 is a point).

\subsection{Warm-up}

Recall that the \emph{unit sphere} in $\mathbf{R}^3$ is the sphere of
radius 1 centered at $(0,0,0)$. Its equation is
\[ x^2 + y^2 + z^2 = 1. \]

\begin{questions}

\question Give equations of all the planes parallel to the $(y,z)$-plane that miss the unit sphere. \emph{Hint.} Use an auxiliary variable.

\flexskip{1}

\question How many planes parallel to the $(y,z)$-plane meet the sphere at one point? Give equations for all of them.

\flexskip{1}
 
\end{questions}

% \subsection{Brute force}

% You should now be pretty convinced that the intersection of a sphere and a
% plane is a circle (provided the intersection contains more than one point). It
% certainly seems \emph{plausible} enough. But is it completely, unambiguously
% obvious that it's impossible to get an elliptical cross-section?

% A completely coordinatized approach to this problem involves choosing a
% sphere with arbitrary center and radius and an arbitrary plane. Using a
% result about the classification of planes (we shall do this after we
% have more vector technology ready to deploy), this will involve solving
% a system of two simultaneous nonlinear equation in eight variables:

% \begin{align*}
%     (x - a)^2 + (y - b)^2 + (z - c)^2 &= r^2 \\
%     Ax + By + Cz &= D 
% \end{align*}

% This approach is easy to set up, but a pain in the rear to work out the
% details, and not very illuminating. Instead, we'll unleash the kung fu
% of symmetry to deal with this complicated geometric situation.

% \subsection{Symmetry and WLOG}

% The idea behind a symmetry argument is that a special case might not
% really be all that special. For us, the special case will be:

% \begin{quote}
% The sphere is the unit sphere and the plane is of the form $z = A$.
% \end{quote}

% It seems very special: after all, there are many, many spheres other
% than the unit sphere, and many planes that are not horizontal. But think
% of it like a videographer: the right angle and zoom turns any situation
% into this special one.

% That's what we mean by symmetry, or the mathematical abbreviation WLOG.
% This stands for \textbf{W}ithout \textbf{L}oss \textbf{O}f
% \textbf{G}enerality. The phrase signifies the idea encapsulated above:
% that what appears to be a special case is in fact sufficiently general.

\subsection{The unit sphere and a horizontal plane}

\begin{questions}

\question Draw a big, beautiful picture of the unit sphere sliced by a horizontal plane. Remember the drawing tips:
\begin{compactitem}
    \item Get the angles right. Make the parallelogram of your plane parallel to the $x$- and $y$-axes.
    \item Draw a \emph{big} picture. Big pictures have more room for labels.
    \item The $(x,y)$-plane itself does slice the sphere, but it's very special. Pick a more generic slicing plane.
    \item Don't be afraid to erase.
\end{compactitem}

\flexskip{5}

\newpage

\question Now, we're still solving two equations. What are they? (Hint: one from the sphere and one from the slicing plane.)

\flexskip{1}

\question Explain why this situation is algebraically preferable to the brute-force solution outlined above.

\flexskip{1}

\question Write down two equations that all the points in the intersection of the sphere and the plane satisfy. (Hint: one equation involves $x$ and $y$ only, and the other involves $z$ only.) These aren't the same as the equations above, but they aren't too terribly different either.

\flexskip{1}

\question Identify the flaw in this attempted use of WLOG.

\begin{quote}
    {\em Consider the cylinder of radius 1 with axis the $x$-axis and the $(y,z)$-plane. The intersection is evidently the circle of radius 1, centered at $(0,0,0)$, and contained in the $(y,z)$-plane. 

    Therefore, we can say that WLOG, whenever a cylinder is cut by a plane, the intersection is a circle. This is because we can always ``rotate and zoom'' to place the cylinder so that its radius is 1 and its axis is the $x$-axis.}
\end{quote}

\flexskip{1}

\end{questions} 

\end{document}