% Module 03:
  Slicing graphs; vectors as displacements
% Dr. Dave Rosoff
% Fall 2013

Last time, we discussed surfaces in $\mathbf{R}^3$: in particular, the graphs of 2-variable functions, spheres, and cylinders. We conclude that discussion with the topic of slicing graphs.

### Slicing graphs

The idea of "slicing graphs" is one that is probably familiar to you, but with disguised terminology. 

#### The "vertical line test" revisited

"The vertical line test" is a criterion used to decide whether a plane curve is the graph of a function (of the horizontal coordinate). To employ it, you visualize all the vertical lines in the plane. Each of these gives a "slice" of the curve---the intersection of the vertical line with the curve. The vertical line test says that the curve is the graph of a function if, and only if, no slice has more than one point in it.

As soon as you find a slice with two (or more) points in it, then you know the curve is not the graph of any function of the horizontal coordinate. 

Slicing curves in the plane is not so interesting, because most of the time the slices are boring: a handful of points, possibly empty. Think of the vertical lines as the knives doing the slicing. In the cases of interest to us, the knives won't necessarily be lines. Instead, they will be planes.

The knives don't have to be vertical for slicing to make sense. It would be possible to talk about the slice obtained by intersecting the curve with any line, of any slope.

#### Slicing in 3 dimensions

3 dimensions is one up from 2, and *surfaces* are one up from *curves*. So what does it mean to slice a surface? We have to say what the knife is. Usually, it will be some plane. The *slice* we get is the intersection of that plane with the surface---usually, a handful of curves.

> **Exercise**. What are the slices that can result from slicing a sphere?
Make sure you include all the possibilities for what the intersection looks like.

> **Exercise**. What does the slice of a circular cylinder look like if the slicing plane is perpendicular to the cylinder? If it is parallel? If it is neither? What can you say about the shapes that are so obtained?

#### The vertical line test for 2-variable functions

Slicing can be used to determine whether a particular surface is the graph of a function of the two horizontal coordinates. 

> **Exercise**. Describe the knives used to slice for the 3-dimensional vertical line test.

> **Exercise**. Formulate the vertical line test for 2-variable functions.

#### Coordinate slices and level curves

The most important special case of slicing surfaces is when the knife is a plane parallel to one of the coordinate planes. The case of the plane $z = 0$ (that is, the $(x,y)$-plane) merits special attention.

> **Definition**. If $f(x,y)$ is a 2-variable function, the slice of its graph obtained by slicing with the plane $z = c$ is called the level curve of $f(x,y)$ with level $c$. In general, it is actually a handful of curves, but the usage does not reflect this. We're stuck with it.

### Displacement numbers

When we are considering motion that is confined to a single line, one number suffices to describe a movement---that is, after we have agreed on the units of the length dimension and the origin of the line. Once these are settled, every pair of positions gives rise to a number: the distance between them. Conventionally, the distance between two positions on a line is always a nonnegative number, but it is convenient to use our notion of negativity as well. This leads to the notion of displacement, which we might also call "directed distance" or "signed distance". The idea is that we no longer regard the two positions as equivalent. Instead, one is the "initial" and one is the "final" position. We agree, once and for all, that displacements are calculated by subtracting the initial position from the final one:

    displacement = pos_final - pos_initial

$$ \Delta x = x_1 - x_0 $$

With these ideas firmly in place, it is not too hard to generalize them to the case of motion in a plane. This is done in the next section. We will assume that the origin and suitable coordinate axes and units are already agreed upon.

Notice that displacement numbers don't depend on a particular starting (or ending) position. The $3$ you get by moving from $2$ to $5$ is identical to the one you get moving from $-166$ to $-163$.

But what distinguishes the *position* $3$ from the *displacement* $3$?

### Displacement vectors

The framework of vectors arises from the desire to be able to compute displacements by subtracting points in the plane, just like in the one- dimensional case described above.

In the one-dimensional case, positions and displacements seem like the same kind of object. Only context allows us to really determine whether a particular $3$ is a position or a displacement between positions. Keep this in mind as you continue reading.

Here is a plot of a few paths connecting the points $(1,1)$ and $(2,3)$ in the plane.

```{r, first_plot, fig.align='center'}
plot(c(1,3), c(1,3), xlab="x", ylab="y")
arrows(1, 1, 2, 2, angle=30, col="red" )
arrows(2, 2, 2, 3, angle=30, col="red" )
arrows(1, 1, 2, 3, angle=30, col="blue")
```

A moment's reflection shows that, no matter how you walk from $(1,1)$ to $(2,3)$, your net change in the $x$-direction is $1$. In the $y$-direction, it is evidently $2$. If we start out at $(2,3)$ and end at $(1,1)$, these changes become negative.

The idea of a displacement *vector* is to bind these two separate numbers, which we might call $\Delta x$ and $\Delta y$, together into one mathematical object. We say the displacement from initial position $(1,1)$ to final position $(2,3)$ is $\langle 1, 2 \rangle$.

> Just like a point in the plane, a displacement vector is specified as a pair of numbers. We distinguish between the two by using round brackets for points $(3,-2)$ and angle brackets for vectors $\langle 3, -2 \rangle$. If we don't want to refer to the coordinates separately, we give the vector a letter name and use an arrow decoration, like $\vec{v}$.

For any particular initial and final point, there's only one displacement between them. But there are many paths between them, which we can visualize as chains of vectors. For example, in the figure above, the points $(1,1)$ and $(2,3)$ are linked via the intermediate point $(2,2)$.

> What are the displacements incurred with initial point $(1,1)$ and final point $(2,2)$? With initial point $(2,2)$ and final point $(2,3)$?

Notice that these displacements add appropriately! Their "sum", adding coordinate-wise, is the overall displacement incurred in moving from $(1,1)$ to $(2,3)$. For this reason, we define the *sum* of two vectors to be the vector whose entries are the sums of the corresponding entries:

$$ \langle x_1, y_1 \rangle + \langle x_2, y_2 \rangle =
     \langle x_1 + x_2, y_1 + y_2 \rangle $$

Observe that the *zero vector* $\langle 0, 0 \rangle$ behaves appropriately under this conception. If adding vectors corresponds to accumulating displacements, adding $\langle 0, 0 \rangle$ should be an operation with no effect---and indeed, it is.

What about subtracting vectors? Well, if we make sense of the *negative* of a vector we will be free to regard $\vec{v} - \vec{w}$ as an abbreviation for $\vec{v} + (-\vec{w})$. A negative displacement has a clear physical meaning: the same distance, but the opposite direction. Therefore negation and subtraction are accomplished coordinate-wise, just like addition.

### Parallel vectors: a simple test

Much of the time we are concerned just with the *direction* of a motion, and not as much with its distance ("magnitude"). We'll need a test to see whether two vectors, described in coordinates, point in the same direction in the plane. This is the subject of [Workshop 02][Workshop 02]. For now, let us discuss the subtleties of the concept of "parallel" in our new context of vectors.

Since lines in the plane are not directed, we don't need a word to describe this situation in ordinary geometry:

```{r, second_plot, fig.align='center'}
plot(c(1.5,3),c(1.5,3),xlab="x",ylab="y")
arrows(1.75,1.5,2.75,2.5, angle=30, col="red")
arrows(3,3,2,2, angle=30, col="blue")
```

Are the vectors parallel, or not? Reasonable people can disagree, and in fact there is not universal agreement on the terminology. We will establish the convention that vectors as pictured above are indeed parallel. Some authors reserve the term "anti-parallel" for this situation.

### Linear relations among vectors

Suppose that $\vec{u}$, $\vec{v}$, $\vec{w}$, and $\vec{z}$ are vectors in the plane. What kind of equations involving them might we consider? We haven't defined most familiar algebraic operations, like multiplication, logarithms, exponents... seems like all we can build are linear equations, like this one:

$$ 2 \vec{u} + 5 \vec{v} = -3 \vec{w} + \vec{z} $$

Regarded as an equation of displacements, it is easy to interpret: the equation states that the overall displacement incurred by following $\vec{u}$ twice and $\vec{v}$ five times is equal to that incurred by following $\vec{w}$ backwards three times and $\vec{z}$ once (forward).

Any such linear equation can be converted, by the standard rules of algebra, into an equation with the zero vector on the right-hand side. Such an equation states that a particular combination of displacements cancels out exactly, that the overall $x$- and $y$-displacements are both $0$.


<!--document ends; links past here-->

[Module 01]: ../01/Module.html
[Module 02]: ../02/Module.html
<!--placeholder link-->
[Workshop 02]: ../../workshops/02/Workshop.pdf 


